---
title: "Reproducible Research: Peer Assessment 1"
output: 
        html_document:
                keep_md: TRUE
---

## Loading and preprocessing the data     
```{r Loading Data setoptions, echo=TRUE}
if(!file.exists('./movement_data.zip')){
        url = 'https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip'
        download.file(url = url, destfile = './movement_data.zip')
        unzip('./movement_data.zip', exdir = './data')
}
activity_dta <- read.csv('./data/activity.csv')
```

Now that we have read the data into our environment let's preprocess it
```{r Preprocessing Data}
library(lubridate)
activity_dta$date <- ymd(activity_dta$date)
```


## What is mean total number of steps taken per day?
```{r Mean Total Steps / Day}
library(dplyr)
# dta_mean <- activity_dta[complete.cases(activity_dta),]
act_dta_day <- 
        activity_dta %>% 
        filter(!is.na(steps)) %>%
        group_by(date) %>% 
        summarise(total_steps = sum(steps, na.rm=TRUE))
hist(act_dta_day$total_steps, xlab = 'Total Steps per Day', main = 'Historgram of Total Steps per Day', col='steelblue')
mean_steps = mean(act_dta_day$total_steps, na.rm=TRUE)
median_steps = median(act_dta_day$total_steps, na.rm = TRUE)
```
Mean steps taken per day is `r mean_steps` steps, while median steps taken per day`r median_steps` steps.

## What is the average daily activity pattern?
```{r}
daily_act_dta <- 
        activity_dta %>%
        filter(!is.na(steps)) %>%
        group_by(interval) %>%
        summarise(avg_steps_interval = mean(steps, na.rm = TRUE))
with(daily_act_dta, plot(interval, avg_steps_interval, type='l', pch=19, xlab='5 min Intervals', ylab='Average Steps', main='Avg Steps per Interval',col='steelblue'))
max_int = daily_act_dta[which.max(daily_act_dta$avg_steps_interval),]
points(max_int, col='red', pch=16)
legend('topright', col=c('steelblue', 'red'), lty=c(1, NA), pch=c(NA,16), legend=c('average steps per 5 min interval', 'interval with max steps'))
```
Maximum number of steps, on average across all the days in the dataset, contains `r max_int$interval`th 5 minute interval.

## Imputing missing values
```{r}
total_nas <- sum(colSums(is.na(activity_dta)))

imp_act_dta <- 
        activity_dta %>%
        group_by(interval) %>%
        mutate(steps = ifelse(is.na(steps), mean(steps,na.rm=TRUE),steps))
imp_act_dta <- as.data.frame(imp_act_dta)

group_imp_day_act_dta <-
        imp_act_dta %>%
        group_by(date) %>%
        summarise(total_steps = sum(steps, na.rm = TRUE))

hist(group_imp_day_act_dta$total_steps, xlab = 'Total Steps per Day', main = 'Historgram of Total Steps per Day (Imputed)', col='steelblue')

imp_mean_steps = mean(group_imp_day_act_dta$total_steps, na.rm = TRUE)
imp_median_steps = median(group_imp_day_act_dta$total_steps, na.rm = TRUE)
```
Total number of missing values in the data set is `r total_nas`.  

Strategy for imputing is to use the mean for each interval group.

The mean total number of steps taken per day did not changed, since we imputed the missing values with mean for each group: from 'r mean_steps' to `r imp_mean_steps'. Median changes slightly: from 'r median_steps` to `r imp+median_steps`.

## Are there differences in activity patterns between weekdays and weekends?
```{r}


```


